
## Here include some basic information
## Name of author:Robert Wynne
## Date:24/01/2020
## Program name: .Rmd
## Purpose of the script: plot euler
## This script does XX, YYY and ZZZ.
##
## Mention what it takes as input
##
## It outputs... - We can work on this when I get back.

First step, load library required for plotting:

```{r, message = FALSE}
##install.packages("eulerr")
library(eulerr)
library(ggplot2)
```

Next, load input data:

```{r, message = FALSE}
load(file = "DESeq2_output_liver.RData")
```

Create a list of comparisons to plot:

```{r, message = FALSE}
vendiagram <- euler(combinations = list("Phenotype genes" = phenotype_contr_env_sig_genes,
                                        "Osmotic environment genes" = env_contr_phen_sig_genes),
                                        shape = "ellipse")
```

Plot Venn diagram

Select all code and run together at the same time to save plot.
It won't work it execute lines of code indepedently.

```{r, message = FALSE}
pdf(file = "results/euler_latest_liver_plot.pdf",
     width = 6,
     height = 4)
plot(vendiagram,
     quantities = TRUE,
     edges = TRUE)
dev.off()
```

For now, save everything:

```{r, message = FALSE}
save.image(file = "eulerr_output_liver.RData")
```