
## Here include some basic information
## Name of author:Robert Wynne
## Date:24/01/2020
## Program name: .Rmd
## Purpose of the script: plot euler
## This script does XX, YYY and ZZZ.
##
## Mention what it takes as input
##
## It outputs... - We can work on this when I get back.

First step, load library required for plotting:

```{r, message = FALSE}
##install.packages("eulerr")
library(eulerr)
library(ggplot2)
```

Next, load input data:

```{r, message = FALSE}
load(file = "DESeq2_output_liver.RData")
```

Create a list of comparisons to plot:

```{r, message = FALSE}
vendiagram <- euler(combinations = list("Smolt-SW vs Resident-SW" = smolt_vs_non_smolts_salt_lrt_wald,
                                        "Smolt-SW vs Resident-FW" = smolt_vs_non_smolts_fresh_lrt_wald,
                                        "Resident-SW vs Resident-FW" = non_smolt_salt_vs_non_smolts_fresh_lrt_wald))
```

Plot Venn diagram

Select all code and run together at the same time to save plot.
It won't work it execute lines of code indepedently.

```{r, message = FALSE}
pdf(file = "results/euler_plot.pdf",
     width = 6,
     height = 4)
plot(vendiagram,
     quantities = TRUE,
     edges = TRUE)
dev.off()
```

Extract genes identified in all three comparisons:

```{r, message = FALSE}
combined_gene_list <- c(smolt_vs_non_smolts_salt_lrt_wald,
                        smolt_vs_non_smolts_fresh_lrt_wald,
                        non_smolt_salt_vs_non_smolts_fresh_lrt_wald)

## Count incidences of each gene:
combined_gene_table <- as.data.frame(cbind(table(combined_gene_list)))

## Subset DEGs occurring in all three comparisons:
three_group_degs <- subset(combined_gene_table,
                           V1 == 3)
nrow(three_group_degs)
## Create gene-list:
three_group_gene_list <- row.names(three_group_degs)
```

Extract genes identified as unique for Smolt-SW vs Resident-SW:

```{r, message = FALSE}
## Extract DEGs unique to smolt_sw vs resident_sw:
smolt_sw_vs_res_sw_degs <- subset(combined_gene_table,
                                  V1 == 1 &
                                          row.names(combined_gene_table) %in%
                                          smolt_vs_non_smolts_salt_lrt_wald)
## Create gene_list:
smolt_sw_vs_res_sw_gene_list <- row.names(smolt_sw_vs_res_sw_degs)

## Extract DEGs unique to smolt_sw vs resident_fw:
smolt_sw_vs_res_fw_degs <- subset(combined_gene_table,
                                  V1 == 1 &
                                          row.names(combined_gene_table) %in%
                                          smolt_vs_non_smolts_fresh_lrt_wald)
## Create gene_list:
smolt_sw_vs_res_fw_gene_list <- row.names(smolt_sw_vs_res_fw_degs)
```

For now, save everything:

```{r, message = FALSE}
save.image(file = "eulerr_output_liver.RData")
```
